Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow

### 2014-07-23 writeups/dissertation/dissertation.tex
foo

The entire StackOverflow dataset is packaged and released quarterly to the public.
So this package was downloaded and the database tables were imported into PostgreSQL using \textcite{so2db} importing tool.

@@

Twitter provides two types of APIs for acquiring data: A query-based and streaming interface.
The \textcite{Tweepy} Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
The query-based interface was used to collect all tweets for specific popular users.
At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.

@@

The base-level ($B_{i}$) component in Table \ref{ACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{ModACTRModel}).
\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk is likely to be presented again, given past frequency.
If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
Both forms maintain the same rank ordering, so they should produce similar results.
So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).

@@

To better illustrate the dynamics of each user's tagging behavior,
profile plots for a single user for both StackOverflow and Twitter are included in Figures \ref{figHTByTime-520957} and \ref{figHTByTime-fashionista_com} respectively.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTByTime-520957-crop.pdf}}}
  \caption{Tagging profile of a single user on StackOverflow}
  \label{figHTByTime-520957}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTByTime-fashionista_com-crop.pdf}}}
  \caption{Tagging profile of a single user on Twitter}
  \label{figHTByTime-fashionista_com}
\end{figure}

@@

As a first method for understanding model performance, profile plots of model's predicted tags for individual users were overlaid on that user's actual tag use.
These plots for StackOverflow and Twitter are included in Figures \ref{figHTMByTime-520957} and \ref{figHTMByTime-fashionista_com} respectively.
These are the same plots shown in Figures \ref{figHTByTime-520957} and \ref{figHTByTime-fashionista_com} with model predictions overlaid in red.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
  \caption{Model tagging profile of a single user on StackOverflow}
  \label{figHTMByTime-520957}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
  \caption{Model tagging profile of a single user on Twitter}
  \label{figHTMByTime-fashionista_com}
\end{figure}

@@

Moving up one level of analysis, model accuracy for all users in a single dataset slice was analyzed.
A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDs} and \ref{figPriorTwitterSliceDs} respectively.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}}
  \caption{Model performance for a single dataset slice for StackOverflow}
  \label{figPriorSOQSliceDsStd}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}}
  \caption{Model performance for a single dataset slice for StackOverflow}
  \label{figPriorSOQSliceDsOL}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}}
  \caption{Model performance for a single dataset slice for Twitter}
  \label{figPriorTwitterSliceDsStd}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}}
  \caption{Model performance for a single dataset slice for Twitter}
  \label{figPriorTwitterSliceDsOL}
\end{figure}

@@

To identify the best decay rate value for the dataset slices,
each user's best decay rate value was taken and the results are plotted in histogram form in Figures \ref{figPriorSOQHistDs} and \ref{figPriorTwitterSliceDs} respectively.

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}}
  \caption{Best-fit decay rate for a single dataset slice for StackOverflow}
  \label{figPriorSOQHistDs}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}}
  \caption{Best-fit decay rate for a single dataset slice for StackOverflow}
  \label{figPriorSOQHistDsOL}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}}
  \caption{Best-fit decay rate for a single dataset slice for Twitter}
  \label{figPriorTwitterHistDs}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{visHistD-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}}
  \caption{Best-fit decay rate for a single dataset slice for Twitter}
  \label{figPriorTwitterHistDsOL}
\end{figure}

@@

The impact of adding the offset to each of the models for StackOverflow and Twitter are included in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetTwitter}.

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTPerm-crop.pdf}}}
  \caption{Stop-word techniques for the random-permutation model for Twitter}
  \label{figContextStopRPSO}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTSji-crop.pdf}}}
  \caption{Stop-word techniques for the Bayesian model for Twitter}
  \label{figContextStopBSO}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
  \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
  \label{figContextLogoddsT}
\end{figure}

@@

Model accuracy when using a user-customized co-occurrence matrix for StackOverflow and Twitter is shown in Figures \ref{figPUserCustomSO} and \ref{figPUserCustomT}.

>>>>

The entire StackOverflow dataset is packaged and released quarterly to the public.
So this package was downloaded and the database tables were imported into PostgreSQL.

@@

Twitter provides two types of APIs for acquiring data: A query-based and streaming interface.
The Tweepy Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
The query-based interface was used to collect all tweets for specific popular users.
At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.

@@

The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk is likely to be presented again, given past frequency.
If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
Both forms maintain the same rank ordering, so they should produce similar results.
So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).

@@

To better illustrate the dynamics of each user's tagging behavior,
profile plots for a single user for both StackOverflow and Twitter are included in Figure \ref{figHTByTime}.

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][StackOverflow]{\includegraphics[width=8cm]{HTByTime-520957-crop.pdf}}
    \hfill
    \subfigure[][Twitter]{\includegraphics[width=8cm]{HTByTime-fashionista_com-crop.pdf}}
    \caption{Tagging profile for a single user}
    \label{figHTByTime}
  }%
\end{figure}

@@

As a first method for understanding model performance, profile plots of model's predicted tags for individual users were overlaid on that user's actual tag use.
These plots for StackOverflow and Twitter are included in Figures \ref{figHTMByTime-520957} and \ref{figHTMByTime-fashionista_com} respectively.
These are the same plots shown in Figure \ref{figHTByTime} with model predictions overlaid in red.

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-520957-crop.pdf}}}
  \caption{Model tagging profile of a single user on StackOverflow}
  \label{figHTMByTime-520957}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{HTMByTime-fashionista_com-crop.pdf}}}
  \caption{Model tagging profile of a single user on Twitter}
  \label{figHTMByTime-fashionista_com}
\end{figure}

@@

Moving up one level of analysis, model accuracy for all users in a single dataset slice was analyzed.
A plot of the results for one of the 12 slices, for both Twitter and StackOverflow, and for both forms of $B_{i}$ are included in Figures \ref{figPriorSOQSliceDsStd} and \ref{figPriorTwitterSliceDsStd} respectively.

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
    \hfill
    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visNormMean-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
    \caption{Model performance for a single dataset slice for StackOverflow}
    \label{figPriorSOQSliceDsStd}
  }%
\end{figure}

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][Standard form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
    \hfill
    \subfigure[][Optmized learning form]{\includegraphics[width=8cm]{visNormMean-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
    \caption{Model performance for a single dataset slice for Twitter}
    \label{figPriorTwitterSliceDsStd}
  }%
\end{figure}

@@

To identify the best decay rate value for the dataset slices,
each user's best decay rate value was taken and the results are plotted in histogram form in Figures \ref{figPriorSOQHistDsOL} and \ref{figPriorTwitterHistDsOL} respectively.

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][Standard form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorStd-crop.pdf}}
    \hfill
    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-SOQgt500r2-topHashtagPostPriorOL2-crop.pdf}}
    \caption{Best-fit decay rate for a single dataset slice for StackOverflow}
    \label{figPriorSOQHistDsOL}
  }%
\end{figure}

\begin{figure}[!htbp]
  {%
    \setlength{\fboxsep}{0pt}%
    \setlength{\fboxrule}{1pt}%
    \hfill
    \subfigure[][Standard form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorStd-crop.pdf}}
    \hfill
    \subfigure[][Optimized learning form]{\includegraphics[width=8cm]{visHistD-TFollowgt10Mr2-topHashtagPostPriorOL2-crop.pdf}}
    \caption{Best-fit decay rate for a single dataset slice for Twitter}
    \label{figPriorTwitterHistDsOL}
  }%
\end{figure}

@@

The impact of adding the offset to each of the models for StackOverflow and Twitter are included in Figures \ref{figContextOffsetSO} and \ref{figContextOffsetT}.

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTPerm-crop.pdf}}}
  \caption{Stop-word techniques for the random-permutation model for Twitter}
  \label{figContextStopRPT}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-allWeightingsTSji-crop.pdf}}}
  \caption{Stop-word techniques for the Bayesian model for Twitter}
  \label{figContextStopBT}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-orderT-crop.pdf}}}
  \caption{Model accuracy for various word order methods for the random-permutation model for Twitter}
  \label{figContextOrder}
\end{figure}

@@

Model accuracy when using a user-customized co-occurrence matrix for StackOverflow and Twitter is shown in Figures \ref{figPUCustomSO} and \ref{figPUCustomT}.
